A Framework for Equivalence of Cast Calculi and a new Cast Representation

### 2020-10-09 paper.tex
Updates from Overleaf
minor edits

% operators
\newcommand{\mbind}[0]{\ensuremath{\;\gg=\;}}
\newcommand{\mcomp}[0]{\ensuremath{\;>=>\;}}

@@

Of course, a space-efficient cast representation must be proved
correct. That is, a language implementation that uses space-efficient
casts should give the same output as the language semantics. Our
generic CEK machine comes with generic theorems mechanized in
the Agda proof assistant~\citep{norell2008dependently}
(Section~\ref{sec:framework}) that researchers can use to prove a cast
representation correct.  We believe this framework can reduce the
burden of proving correctness theorems and help people focus on the
cast representations. As a demonstration, we apply the framework to
Lazy UD coercions in normal form~\citep{Siek:2015:BCT:2737924.2737968}
and Lazy UD threesomes~\citep{Siek:2010:TWB:1706299.1706342}. We
believe the framework is also applicable to other cast
representations. The framework and demonstrations are available at the
following URL:
\[\text{\urlFrameworkAndIllustration}\]

@@

The syntax \error{l} occurs several times in the syntax definition. We shall 
view those nonterminals that have an \error{l} case as error monads. We use 
the symbol \mbind{} as the monad bind operator, which is defined as follows:
\[
\begin{array}{rcll}
(\error{l}) \mbind f & = & \error{l} \\
          a \mbind f & = & f(a)      & \text{otherwise}
\end{array}
\]
\todo[inline]{JS: why not use monadic let instead? That avoids
  the need to also use meta-lambdas, which you would need to
  explain, e.g., in the transition rule that uses applyCast.}
\todo[inline]{KC: 1. I will try to switch to monadic let. As you pointed out
when we wrote the WGT paper, let is more friendly to readers who are not familiar
with monads.
2. Do you imply that monadic let doesn't need to be explained?
}
\todo[inline]{JS: No, you should still define monadic let.}
\todo[inline]{KC: will do.}

@@

The transition relation \reduceto{s}{s} invokes the cast application
function, noted as $\applyCast{v}{\ccast{S}{l}{T}}$, which is also defined
by specify blame strategies (Fig.~\ref{fig:blame-strategy-LazyD} and
Fig.~\ref{fig:blame-strategy-LazyUD}).
\todo[inline]{JS: be consistent with the font used for applyCast. Sometimes
  you use roman and sometimes you used italic. Also, be consistent
  for all the metafunctions (eval, observe, etc.).}
\todo[inline]{KC: Fixed.}
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%
To cast a value, the machine invokes applyCast on the value. The cast
result is either a value or an error (\error{l}). If the cast
succeeds, the machine returns the value to the next continuation.  If
the cast fails, the machine halts with the blame label.
%
To apply a casted function, the machine first casts the argument $v$, then
applies the inner function $u$ to the casted argument, and finally
casts the return value.
%
To project the first element of a casted pair, the machine projects
the first element of the inner pair $u$ and cast it.
%
Projecting the second element of a casted pair is similar.

@@

\todo[inline]{KC: conditions of blame strategy. Jeremy, please }
\begin{definition}[Blame Strategy, $B$]
A blame strategy defines a subset of pre-types, written I, that is injectable,
and a function, written \applyCastName, that applies cast to values.
\end{definition}

>>>>

% operators
\newcommand{\mbind}[0]{\ensuremath{\;\gg=\;}}
\newcommand{\mlet}[3]{\ensuremath{
\mathtt{let}\ #1\leftarrow#2\ \mathtt{in}\;#3}}
\newcommand{\mcomp}[0]{\ensuremath{\;>=>\;}}

@@

Of course, a space-efficient cast representation must be proved
correct. That is, a language implementation that uses space-efficient
casts should give the same output as the language semantics. Our
generic CEK machine comes with generic theorems mechanized in
the Agda proof assistant~\citep{norell2008dependently}
(Section~\ref{sec:framework}) that researchers can use to prove a cast
representation correct.  We believe this framework can reduce the
burden of proving correctness theorems and help people focus on the
cast representations. As a demonstration, we apply the framework to
Lazy UD coercions in normal form~\citep{Siek:2015:BCT:2737924.2737968}
and Lazy UD threesomes~\citep{Siek:2010:TWB:1706299.1706342}. We
believe the framework is also applicable to other cast
representations. The framework and demonstrations are available at the
following URL:
\[\text{\urlFrameworkAndIllustration}\]
\todo[inline]{KC: is this the right place to mention that in Agda
code we use de Bruijn indexes?}

@@

The syntax \error{l} occurs several times in the syntax definition. We shall 
view those nonterminals that have an \error{l} case as error monads.
That is, they raise an error with the error message $l$.
The monadic \texttt{let} is defined as follows:
\[
\begin{array}{rcll}
    (\mlet{x}{(\error{l})}{e}) & = & \error{l} \\
    (\mlet{x}{a}{e}) & = & [a/x]e & \text{otherwise} \\ 
\end{array}
\]
In Agda code we use the monad bind operator instead.
% We use 
% the symbol \mbind{} as the monad bind operator, which is defined as follows:
% \[
% \begin{array}{rcll}
% (\error{l}) \mbind f & = & \error{l} \\
%           a \mbind f & = & f(a)      & \text{otherwise}
% \end{array}
% \]
\todo[inline]{JS: why not use monadic let instead? That avoids
  the need to also use meta-lambdas, which you would need to
  explain, e.g., in the transition rule that uses applyCast.}
\todo[inline]{KC: 1. I will try to switch to monadic let. As you pointed out
when we wrote the WGT paper, let is more friendly to readers who are not familiar
with monads.
2. Do you imply that monadic let doesn't need to be explained?
}
\todo[inline]{JS: No, you should still define monadic let.}
\todo[inline]{KC: KC, you need to replace all uses of monad bind to let.}

@@

The transition relation \reduceto{s}{s} invokes the cast application
function, noted as $\applyCast{v}{\ccast{S}{l}{T}}$, which is also defined
by specify blame strategies (Fig.~\ref{fig:blame-strategy-LazyD} and
Fig.~\ref{fig:blame-strategy-LazyUD}).
\todo[inline]{JS: be consistent with the font used for applyCast. Sometimes
  you use roman and sometimes you used italic. Also, be consistent
  for all the metafunctions (eval, observe, etc.).}
\todo[inline]{KC: Fixed}
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%
To cast a value, the machine invokes applyCast on the value. The cast
result is either a value or an error (\error{l}). If the cast
succeeds, the machine returns the value to the next continuation.  If
the cast fails, the machine halts with the blame label.
%
To apply a casted function, the machine first casts the argument $v$, then
applies the inner function $u$ to the casted argument, and finally
casts the return value.
%
To project the first element of a casted pair, the machine projects
the first element of the inner pair $u$ and cast it.
%
Projecting the second element of a casted pair is similar.

@@

\begin{definition}[Blame Strategy, $B$]
A blame strategy defines a subset of pre-types, written I, that is injectable,
and a function, written \applyCastName, that applies cast to values.
\end{definition}
